---
title: "Ruth's GPS"
output: 
  pdf_document: 
    fig_height: 9
    fig_caption: yes
---
HALTED for now while the extra data are located.


Version 4 triad: like 4, but focuses on the three sensors left close together: GPS Mallemuk, Ismåge and Havtern (+ Strandskade)  that were deployed in a triangle from 28/03 ~~ 19.30 to 30/03 15.00 (UTC)

```{r, message=FALSE,warning=FALSE}
rm(list=ls())
library(MASS)
library(ie2misc)
library(dplyr)
library(lubridate)
setwd("~/WORKSHOP/GPS/")

df <- readRDS("OUTPUT/EVERYTHING3.rds")
# clean out columns with only NA
listen <- NULL
for (icol in 1:ncol(df))
{    listen <- rbind(listen,c(icol,colSums(is.na(df[,icol]))))}
idx <- which(listen[,2] == nrow(df))
df <- df[,-idx]
df$POSIX <- as.POSIXct(df$"Timestamp UTC",tz="UTC")
df_org <- df
#
unique_names <- unique(df$UnitName)
unique_names

```

# Select just the four mentioned
```{r}

# select just the four stations mentioned
fourstations <- c("Mallemuk 300434066431710","Ismaage 300434066437720","Havterne 300434066435700","Strandskade 860640050251356")
dfnew <- NULL
for (iset in fourstations)
     {
      idx <- which(df$UnitName == iset) 
      print(c(iset,length(idx)))
      dfnew <- rbind.data.frame(dfnew,df[idx,])
}
# Snip to the time interval mentioned
idx <- which(dfnew$POSIX >= as.POSIXct("2022-03-28 19:30:00",tz="UTC") & dfnew$POSIX < as.POSIXct("2022-03-30 15:00:00",tz="UTC"))

```


# Define that function for later use
```{r}
kleaner <- function(df_in)
{
  # clean out columns with only NA
  listen <- NULL
  for (icol in 1:ncol(df_in))  {    listen <- rbind(listen,c(icol,colSums(is.na(df_in[,icol]))))}
  
  idx <- which(listen[,2] == nrow(df_in))
  if (length(idx) != 0) {df_in <- df_in[,-idx]}
  return(df_in)
}
```

# need to clean up Fjeldrype as it is the reference
```{r}
par(mfrow=c(2,1))
df <- df_org
idx <- which(df$UnitName == "Fjeldrype 860640050232018")
# remove the old
df_org <- df_org[-idx,]
#
df_here <- df[idx,]
df_here <- na.omit(df_here)
plot(df_here$POSIX,df_here$Longitude,type="b",main="Fjeldrype before outlier removed")
rlmfit <- rlm(df_here$Longitude ~ df_here$POSIX)
z <- unname(abs(rlmfit$residuals/madstat(rlmfit$residuals))) # counting goes off due to omitted values in residuals - they are not NAs, they are simply missing! See, e.g. https://stackoverflow.com/questions/6882709/how-do-i-deal-with-nas-in-residuals-in-a-regression-in-r
jdx <- which(z > 7)
df_here$Longitude[jdx] <- NA
print(paste("I removed in Fjeldrype ",length(jdx)," outliers in longitude"))
plot(df_here$POSIX,df_here$Longitude,type="b",main="Fjeldrype after outlier removed")
# rbind the new
df_org <- rbind.data.frame(df_org,df_here)
df <- df_org
```
# Also clean up Ismaage
```{r}
par(mfrow=c(2,1))
df <- df_org
idx <- which(df$UnitName == "Ismaage 300434066437720")

#
df_here <- df[idx,]
#df_here <- na.omit(df_here) # too general
ldx <- which(is.na(df_here$Longitude))
if (length(ldx) != 0) {df_here <- df_here[-ldx,]}
plot(df_here$POSIX,df_here$Longitude,type="b",main="Ismaage before outlier removed")
rlmfit <- rlm(df_here$Longitude ~ df_here$POSIX)
z <- unname(abs(rlmfit$residuals/madstat(rlmfit$residuals))) # counting goes off due to omitted values in residuals - they are not NAs, they are simply missing! See, e.g. https://stackoverflow.com/questions/6882709/how-do-i-deal-with-nas-in-residuals-in-a-regression-in-r
jdx <- which(z > 7)
df_here$Longitude[jdx] <- NA
print(paste("I removed in Ismaage ",length(jdx)," outliers in longitude"))
plot(df_here$POSIX,df_here$Longitude,type="b",main="Ismaage after outlier removed")
# remove the old
df_org <- df_org[-idx,]
# rbind the new
df_org <- rbind.data.frame(df_org,df_here)
df <- df_org
```
# Also clean up Mallemuk lon
```{r}
par(mfrow=c(2,1))
df <- df_org
idx <- which(df$UnitName == "Mallemuk 300434066431710")

#
df_here <- df[idx,]
#df_here <- na.omit(df_here) # too general
ldx <- which(is.na(df_here$Longitude))
if (length(ldx) != 0) {df_here <- df_here[-ldx,]}
plot(df_here$POSIX,df_here$Longitude,type="b",main="Mallemuk before outlier removed")
rlmfit <- rlm(df_here$Longitude ~ df_here$POSIX)
z <- unname(abs(rlmfit$residuals/madstat(rlmfit$residuals))) # counting goes off due to omitted values in residuals - they are not NAs, they are simply missing! See, e.g. https://stackoverflow.com/questions/6882709/how-do-i-deal-with-nas-in-residuals-in-a-regression-in-r
jdx <- which(z > 4)
df_here$Longitude[jdx] <- NA
print(paste("I removed in Mallemuk ",length(jdx)," outliers in longitude"))
plot(df_here$POSIX,df_here$Longitude,type="b",main="Mallemuk after outlier removed")
# remove the old
df_org <- df_org[-idx,]
# rbind the new
df_org <- rbind.data.frame(df_org,df_here)
df <- df_org
```

# Also clean up Mallemuk lat
```{r}
par(mfrow=c(2,1))
df <- df_org
idx <- which(df$UnitName == "Mallemuk 300434066431710")

#
df_here <- df[idx,]
#df_here <- na.omit(df_here) # too general
ldx <- which(is.na(df_here$Longitude))
if (length(ldx) != 0) {df_here <- df_here[-ldx,]}
plot(df_here$POSIX,df_here$Latitude,type="b",main="Mallemuk before outlier removed")
rlmfit <- rlm(df_here$Latitude ~ df_here$POSIX)
z <- unname(abs(rlmfit$residuals/madstat(rlmfit$residuals))) # counting goes off due to omitted values in residuals - they are not NAs, they are simply missing! See, e.g. https://stackoverflow.com/questions/6882709/how-do-i-deal-with-nas-in-residuals-in-a-regression-in-r
jdx <- which(z > 4)
df_here$Latitude[jdx] <- NA
print(paste("I removed in Mallemuk ",length(jdx)," outliers in Latitude"))
plot(df_here$POSIX,df_here$Latitude,type="b",main="Mallemuk after outlier removed")
# remove the old
df_org <- df_org[-idx,]
# rbind the new
df_org <- rbind.data.frame(df_org,df_here)
df <- df_org
```


## Utility GC formula
```{r}

# Calculates the geodesic distance between two points specified by radian latitude/longitude using the
# Haversine formula (hf)
gcd.hf <- function(long1, lat1, long2, lat2) {
  R <- 6371 # Earth mean radius [km]
  delta.long <- (long2 - long1)
  delta.lat <- (lat2 - lat1)
  a <- sin(delta.lat/2)^2 + cos(lat1) * cos(lat2) * sin(delta.long/2)^2
  c <- 2 * asin(min(1,sqrt(a)))
  d = R * c
  return(d) # Distance in km
}

```
# Define function to calculate speed
```{r}
getSpeed <- function(time,lon,lat)
{
  rtod <- pi/180
  speed <- NULL
  for (it in 1:(length(time)-1))
  {
    # calc great-circle distance between pairs of points
    distance <- gcd.hf(rtod*lon[it+1],rtod*lat[it+1],rtod*lon[it],rtod*lat[it])
    delta_time <- as.numeric(time[it+1]-time[it])/60 # dt in hours now
    #browser()
    # calc speed
    speed <- c(speed,abs(distance/delta_time))
  }
  return(list("speed"=speed))
}
```

# read
```{r}
statdat <- NULL
alldf <- NULL
ic <- 1
for (istat in unique_names)
{
  par(mfrow=c(3,3))
  idx <- which(df_org$UnitName == istat & df_org$Longitude < -65)
  df2 <- df_org[idx,]  
  df2 <- kleaner(df2)
  cnams <- colnames(df2)
  time <- as.POSIXct(df2$"Timestamp UTC",tz="UTC")
  #
  lon <- df2$Longitude
  lat <- df2$Latitude
  temperature <- df2$"Temperature(°C)" 
  acceleration <- sqrt(df2$"AccelerationX(g)" ^2+df2$"AccelerationY(g)" ^2+df2$"AccelerationZ(g)" ^2)
  lightlevel <- df2$"LightLevel"
  speed0 <- df2$"GPS Speed(Km/h)"
  speed <- getSpeed(time,lon,lat)$speed
  speed <- c(speed,NA) # need one more value at the end
  
  plot(lon,lat,main=istat,pch=19,cex=0.2,type="p")
  plot(time,lon,main=istat,pch=19,cex=0.2,type="b")
  plot(time,lat,main=istat,pch=19,cex=0.2,type="b")
  if (length(temperature > 3)) {plot(time,temperature,main=istat,pch=19,cex=0.2,type="b")}
  if (length(speed > 3)) {plot(time,speed,main=istat,pch=19,cex=0.2,type="b",ylim=c(0.01,5),log="y")}
  if (length(acceleration > 3)) {
    plot(time,acceleration,main=istat,pch=19,cex=0.2,type="b")
    abline(h=1,col=2,lwd=3)
  }
  if (length(lightlevel > 3)) {plot(time,lightlevel,main=istat,pch=19,cex=0.2,type="b")}
  # the set_of_variables 
  set <- c("time","lon","lat","temperature","acceleration","leightlevel","speed")
  df3 <- cbind.data.frame(time,lon,lat)
  colnames(df3)[1] <- "POSIX"
  if (length(temperature) == nrow(df3)){    df3 <- cbind.data.frame(df3,temperature)  }
  if (length(acceleration) == nrow(df3)){    df3 <- cbind.data.frame(df3,acceleration)  }
  if (length(lightlevel) == nrow(df3)){    df3 <- cbind.data.frame(df3,lightlevel)  }
  if (length(speed) == nrow(df3)){    df3 <- cbind.data.frame(df3,speed)  }
  saveRDS(df3,paste0('OUTPUT/',istat,'.rds'))
  #
  
}

```


# Remove dogsled series
```{r}
if_remove_dogsled <- TRUE
if (if_remove_dogsled){
  unique_names
  idx <- which(unique_names == "Landsvale 860640050251737")
  unique_names <- unique_names[-idx]
  idx <- which(unique_names == "Strandskade 860640050251356")
  unique_names <- unique_names[-idx]
  idx <- which(unique_names == "Ravn 860640050244401")
  unique_names <- unique_names[-idx]
  idx <- which(unique_names == "Stenpikkere 860640050244062")
  unique_names <- unique_names[-idx]
  unique_names
}
```


```{r}
pdf("FIGURES/speed_vs_time.pdf")
# same t-axis
ic <- 1
legtext <- NULL
colnames <- c("green","red","blue","azure","orange","black","purple","salmon","grey","hotpink","yellow","plum")
for (istat in unique_names)
{
  df <- readRDS(paste0('OUTPUT/',istat,'.rds'))
  idx <- order(df$POSIX)
  df <- df[idx,]
  if (ic == 1){   plot(df$POSIX,df$speed, type="b", pch=ic+14,cex=0.8, xlab="Date/Time", ylab="speed [km/hr]", col=colnames[ic],ylim=c(0,1), main="Starting March 28th 2022 at 00 UTC", 
                       xlim=c(as.POSIXct("2022-03-28 00:00:00",tz="UTC"), as.POSIXct("2022-03-31 23:59:59",tz="UTC")))     } 
  else{   lines(df$POSIX,df$speed,col=colnames[ic],type="b",cex=0.8,pch=ic+14)  }
  legtext <- c(legtext,paste(strsplit(istat,' ')[[1]][1],colnames[ic]))
  ic <- ic+1
}
legend("topright",legend=legtext,cex=0.6)
#--------------------------------------------------------------
 
dev.off()
```


# relative to Fjeldrype
```{r,message=FALSE}
par(mfrow=c(1,1))

base_station <- readRDS("OUTPUT/Fjeldrype 860640050232018.rds")

alldf <- NULL
#
for (jstat in 1:length(unique_names))
{
  
  #print(jstat)
  statname <- unique_names[jstat]
  #print(statname)
  png(paste0("FIGURES/offset_lon_lat_rel_Fjeldrype_",statname,".png"))
  
  other <- readRDS(paste0("OUTPUT/",statname,".rds"))
  
  tmin <- max(c(min(base_station$POSIX),min(other$POSIX)))
  tmax <- min(max(base_station$POSIX),max(other$POSIX))
  idx <- which(base_station$POSIX >= tmin & base_station$POSIX <= tmax)
  base_station <- base_station[idx,]
  idx <- which(other$POSIX >= tmin & other$POSIX <= tmax)
  other <- other[idx,]
  #Interpolate to same times as in 'base_station'
  common_t <- base_station$POSIX
  lon_other_interp <- approx(other$POSIX,other$lon,base_station$POSIX,na.rm=TRUE)$y
  lat_other_interp <- approx(other$POSIX,other$lat,base_station$POSIX,na.rm=TRUE)$y
  #
  interp_lon <- na.omit(cbind.data.frame(common_t,lon_other_interp))
  colnames(interp_lon) <- c("POSIX","lon_i")
  interp_lat <- na.omit(cbind.data.frame(common_t,lat_other_interp))
  colnames(interp_lat) <- c("POSIX","lat_i")
  together <- merge(base_station,interp_lon,by="POSIX")
  together <- merge(together,interp_lat,by="POSIX")
  delta_lon <- together$lon_i-together$lon
  delta_lat <- together$lat_i-together$lat
  together <- cbind(together,delta_lon,delta_lat)
  saveRDS(together,paste0("OUTPUT/processed_",statname,".rds"))
  print(paste(statname,round(sd(together$delta_lon),4),round(sd(together$delta_lat),4)))
  plot(together$delta_lon,together$delta_lat,main=statname,xlab="offset lon",ylab="offset lat",pch=19,cex=0.2,type="p",asp=3)
  #
  dev.off()
  plot(together$delta_lon,together$delta_lat,main=statname,xlab="offset lon",ylab="offset lat",pch=19,cex=0.2,type="p",asp=3)
}
```



