---
title: "Plot positions and speeds"
output: 
  pdf_document: 
    fig_height: 9
---


Version 3: calculates speed at the sudden changes in location. This is just an EXAMPLE code acting only on Latitude. Later versions will set up Longitude also as well as the actual ground speed.

Reads .xlsx files
Outputs .rds files for columns not all NA

```{r}
rm(list=ls())
setwd("~/WORKSHOP/GPS/")
library(dplyr)
library(lubridate)
library(MASS)

Rearth <- 6371*1e3 # meters

```

## Utility GC formula
```{r}

# Calculates the geodesic distance between two points specified by radian latitude/longitude using the
# Haversine formula (hf)
gcd.hf <- function(long1, lat1, long2, lat2) {
  R <- 6371*1000 # Earth mean radius [m]
  delta.long <- (long2 - long1)
  delta.lat <- (lat2 - lat1)
  a <- sin(delta.lat/2)^2 + cos(lat1) * cos(lat2) * sin(delta.long/2)^2
  c <- 2 * asin(min(1,sqrt(a)))
  d = R * c
  return(d) # Distance in m
}

```

# Define function to calculate speed
```{r}
getSpeed <- function(time,lon,lat)
{
  rtod <- pi/180
  speed <- NULL
  speed_smoo <- NULL
  for (it in 1:(length(time)-1))
  {
    # calc great-circle distance between pairs of points
    distance <- gcd.hf(rtod*lon[it+1],rtod*lat[it+1],rtod*lon[it],rtod*lat[it])
    delta_time <- as.numeric(time[it+1]-time[it])/60 # dt in hours now
    #browser()
    # calc speed
    speed <- c(speed,abs(distance/delta_time))
  }
  speed_smoo <- NULL
  for (j in 1:(length(speed)-1))
  {
    speed_smoo <- rbind(speed_smoo,c(j,median(c(speed[j-2],speed[j-1],speed[j],speed[j+1],speed[j+1]),na.rm=TRUE)))
  }
  speed_smoo <- rbind(speed_smoo[1,],speed_smoo,speed_smoo[nrow(speed_smoo),])
  #browser()
  return(list("speed"=speed,"speed_smoo"=speed_smoo))
}
```

# Plot coloured points
```{r}
plotcolouredpoints <- function(x,y,limitdates,pair,ivar)
{
  idx <- which(df$POSIX >= limitdates[(pair-1)*2+1] & df$POSIX < limitdates[(pair-1)*2+2])
  points(x[idx],y[idx],type="p",cex=0.3,col=1+pair)
  if (ivar == 'speed') {print(paste(pair,' from ',limitdates[(pair-1)*2+1],' to ', limitdates[(pair-1)*2+2],' speed: ',round(median(y[idx],na.rm=T),1),' +/- ',round(sd(y[idx],na.rm=T)/sqrt(length(idx)),1), 'm/hr'))}
  
}
```



# plot positions and speeds etc
```{r}
plot_stuff <- function(df,name,limitdates)
{
  par(mfrow=c(4,1))
  nlimits <- length(limitdates)
  statname <- strsplit(strsplit(name, "/")[[1]][2],".rds")[[1]][1]
  # First plot positions
  plot(df$Longitude,df$Latitude,type="p",cex=0.3,xlab="lon",ylab="lat",main=statname)
  plotcolouredpoints(df$Longitude,df$Latitude,limitdates,1,'')
  plotcolouredpoints(df$Longitude,df$Latitude,limitdates,2,'')
  plotcolouredpoints(df$Longitude,df$Latitude,limitdates,3,'')
  plotcolouredpoints(df$Longitude,df$Latitude,limitdates,4,'')
  # Plot lon vs time
  plot(df$POSIX,df$Longitude,type="p",cex=0.3,xlab="Date/Time",ylab="lon",main=statname)
  plotcolouredpoints(df$POSIX,df$Longitude,limitdates,1,'')
  plotcolouredpoints(df$POSIX,df$Longitude,limitdates,2,'')
  plotcolouredpoints(df$POSIX,df$Longitude,limitdates,3,'')
  plotcolouredpoints(df$POSIX,df$Longitude,limitdates,4,'')
  # Plot lat vs time
  plot(df$POSIX,df$Latitude,type="p",cex=0.3,xlab="Date/Time",ylab="lat",main=statname)
  plotcolouredpoints(df$POSIX,df$Latitude,limitdates,1,'')
  plotcolouredpoints(df$POSIX,df$Latitude,limitdates,2,'')
  plotcolouredpoints(df$POSIX,df$Latitude,limitdates,3,'')
  plotcolouredpoints(df$POSIX,df$Latitude,limitdates,4,'')
  # get speed
  out <- getSpeed(df$POSIX,df$Longitude,df$Latitude)
  speed <- out$speed
  speed_smoo <- out$speed_smoo[,2]
  df$speed <- c(speed[1],speed)
  df$speed_smoo <- speed_smoo
  #browser()
  # plot speed against time
  plot(df$POSIX,df$speed,type="p",cex=0.3,xlab="Date/Time",ylab="speed [m/hr]",main=statname)
  plotcolouredpoints(df$POSIX,df$speed,limitdates,1,'speed')
  plotcolouredpoints(df$POSIX,df$speed,limitdates,2,'speed')
  plotcolouredpoints(df$POSIX,df$speed,limitdates,3,'speed')
  plotcolouredpoints(df$POSIX,df$speed,limitdates,4,'speed')
  #lines(df$POSIX,df$speed_smoo)
}
```

# model positions and calculate speeds at jumps
```{r}
model_motion <- function(df,name,limitdates)
{
  
  par(mfrow=c(3,1))
  nlimits <- length(limitdates)
  statname <- strsplit(strsplit(name, "/")[[1]][2],".rds")[[1]][1]
  # Latitude
  latitude_pred_at_interval_left_right <- NULL
  lat_speed <- NULL
  # loop over limidates and model positions before and after each limitdate
  for (ilimit in seq(from=1,to=(nlimits-1),by=2))
  {
    idx <- which(df$POSIX >= limitdates[ilimit] & df$POSIX < limitdates[ilimit+1] & !is.na(df$Latitude) )
    rlmfit <- rlm(df$Latitude[idx] ~ df$POSIX[idx])
    lat_speed <- rbind.data.frame(lat_speed,c(ilimit,summary(rlmfit)$coefficients[2]*3600/180*pi)) # radians per hour
    #print(c(rlmfit$fitted.values[1],last(rlmfit$fitted.values)))
    if (ilimit == 1) {
      #browser()
      plot(df$POSIX[idx],df$Latitude[idx],type="p",xlim=range(df$POSIX),ylim=range(df$Latitude,na.rm=T),xlab="Time",ylab="Latitude",main=statname)
      lines(df$POSIX[idx],rlmfit$fitted.values,col=2,lwd=3)
      # evaluate diff at jump
      latitude_pred_at_interval_left_right  <- c(first(rlmfit$fitted.values), last(rlmfit$fitted.values))
    }
    if (ilimit  > 1) {
      points(df$POSIX[idx],df$Latitude[idx])
      lines(df$POSIX[idx],rlmfit$fitted.values,col=2,lwd=3)
      #
      latitude_pred_at_interval_left_right  <- rbind.data.frame(latitude_pred_at_interval_left_right, c(first(rlmfit$fitted.values), last(rlmfit$fitted.values)))
      
    }
    
  }
  colnames(lat_speed) <- c("segment_number","lat_speed_radperhr")
  # Longitude
  longitude_pred_at_interval_left_right <- NULL
  lon_speed <- NULL
  # loop over limidates and model positions before and after each limitdate
  for (ilimit in seq(from=1,to=(nlimits-1),by=2))
  {
    idx <- which(df$POSIX >= limitdates[ilimit] & df$POSIX < limitdates[ilimit+1] & !is.na(df$Longitude) )
    rlmfit <- rlm(df$Longitude[idx] ~ df$POSIX[idx])
    lon_speed <- rbind.data.frame(lon_speed,c(ilimit,summary(rlmfit)$coefficients[2]*3600/180*pi)) # radians per hour
    #print(c(rlmfit$fitted.values[1],last(rlmfit$fitted.values)))
    if (ilimit == 1) {
      #browser()
      plot(df$POSIX[idx],df$Longitude[idx],type="p",xlim=range(df$POSIX),ylim=range(df$Longitude,na.rm=T),xlab="Time",ylab="Longitude",main=statname)
      lines(df$POSIX[idx],rlmfit$fitted.values,col=2,lwd=3)
      # evaluate diff at jump
      longitude_pred_at_interval_left_right  <- c(first(rlmfit$fitted.values), last(rlmfit$fitted.values))
    }
    if (ilimit  > 1) {
      points(df$POSIX[idx],df$Longitude[idx])
      lines(df$POSIX[idx],rlmfit$fitted.values,col=2,lwd=3)
      #
      longitude_pred_at_interval_left_right  <- rbind.data.frame(longitude_pred_at_interval_left_right, c(first(rlmfit$fitted.values), last(rlmfit$fitted.values)))
      
    }
    
  }
  colnames(lon_speed) <- c("segment_number","lon_speed_radperhr")
  segment_speed <- Rearth*sqrt((lon_speed[,2]*cos(median(df$Latitude,na.rm=T)/180*pi))^2+(lat_speed[,2])^2) # meters per hour
  #browser()
  return(list("lats"=latitude_pred_at_interval_left_right,"longs"=longitude_pred_at_interval_left_right,"segspeed"=segment_speed))
}
```

# function to get interval and jump speeds
```{r}
get_jump_speeds <- function(listerne,lon_in,lat_in)
{
  lon <- lon_in/180*pi # in radians
  lat <- lat_in/180*pi
  delta_t <- 3 # hours
  
  # calculate jump speeds
  n_segments <- nrow(listerne$lats)
  speed <- NULL
  for (iseg in 1:(n_segments-1))
  {
    delta_longitude <- listerne$longs[iseg+1,1]-listerne$longs[iseg,2] # in degrees
    delta_longitude <- delta_longitude/180*pi # in radians
    delta_latitude  <- listerne$lats[iseg+1,1]-listerne$lats[iseg,2] # in degrees
    delta_latitude <- delta_latitude/180*pi # in radians
    speed <- rbind.data.frame(speed,c(iseg,Rearth/delta_t*sqrt(delta_longitude^2*cos(lat)^2+delta_latitude^2))) # in meters/hr
  }
  colnames(speed) <- c("jump_number","speed_metersph")
  return(list("speed_jump"=speed))
}
```


# read and plot each file
```{r,fig.width=6,fig.height=9}
files <- c('OUTPUT/Mallemuk.rds','OUTPUT/Soekonge.rds','OUTPUT/Ismaage.rds','OUTPUT/Havterne.rds','OUTPUT/Havoern.rds','OUTPUT/Edder.rds')
#files <- files[1]
# list the important times - start, jumps, ending:
important_times <- c(as.POSIXct("2022-03-31 00:00:00",tz="UTC"),as.POSIXct("2022-04-7 00:00:00",tz="UTC"),
                     as.POSIXct("2022-04-24 12:00:00",tz="UTC"),as.POSIXct("2022-04-27 17:00:00",tz="UTC"),
                     as.POSIXct("2022-05-04 00:00:00",tz="UTC"),as.POSIXct("2022-05-31 00:00:00",tz="UTC"))
limitdates <- NULL
for (it in 1:(length(important_times)-1))
{
  limitdates <- c(limitdates,c(important_times[it],important_times[it+1]))
}

#limitdates2 <- c(as.POSIXct("2022-03-31 00:00:00",tz="UTC"),as.POSIXct("2022-04-7 00:00:00",tz="UTC"),
#                as.POSIXct("2022-04-7 00:00:00",tz="UTC"),as.POSIXct("2022-04-24 12:00:00",tz="UTC"),
#                as.POSIXct("2022-04-24 12:00:00",tz="UTC"),as.POSIXct("2022-04-27 17:00:00",tz="UTC"),
#                as.POSIXct("2022-04-27 17:00:00",tz="UTC"),as.POSIXct("2022-05-04 00:00:00",tz="UTC"),
#                as.POSIXct("2022-05-04 17:00:00",tz="UTC"),as.POSIXct("2022-05-31 00:00:00",tz="UTC")) 
#browser()
for (ifil in 1:length(files))
{
  print("------------------------------------------")
  print(paste(" Processing file ",files[ifil]))
  df <- readRDS(files[ifil])
  plot_stuff(df,files[ifil],limitdates)
  # more better way
  listerne <- model_motion(df,files[ifil],limitdates)
  segspeeds <- round(listerne$segspeed,2)
  print(segspeeds)
  # get speeds of various kinds from the list of jumps
  speeds <- get_jump_speeds(listerne,lon=median(df$Longitude,na.rm=T),lat=median(df$Latitude,na.rm=T))
  print(speeds)
  
}
print("------------------------------------------")
```


